Untitled

### 2021-08-20 analysis/paper/paper.Rmd
combine sections into single document

@@

>>>>

---
title: "Untitled"
author:
  - "Bjørn Peare Bartholdy"
  - Amanda G. Henry
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    df_print: kable
    toc: no
    latex_engine: xelatex
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2: default
  bookdown::html_document2:
    df_print: kable
    keep_md: true
linestretch: 2.0 # double-spacing for review purposes
bibliography: paper/refs.bib
csl: paper/apa.csl
abstract: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.path = "./figures/")
library(tidyverse)
library(broom)
library(here)
library(patchwork)
# load R scripts
source(here("analysis/scripts/analysis.R"))
source(here("analysis/scripts/tables.R"))
source(here("analysis/scripts/figures.R"))
## Compile refs.bib using rbbt package ##
file_paths <- here("analysis/paper", c("intro.Rmd", "matmet.Rmd", "disc.Rmd"))
cite_keys <- sapply(file_paths, rbbt::bbt_detect_citations) 
cite_keys <- unlist(cite_keys, use.names = FALSE)
rbbt::bbt_write_bib("./paper/refs.bib", keys = cite_keys, overwrite = T)
## end refs.bib ##
```

# Introduction

<!-- basic introduction to the field, comprehensible to a scientist in any discipline -->
<!-- Dental calculus is mineralised plaque. And it is pretty cool. -->
Dental calculus, mineralised dental plaque, has proven to contain a wealth of 
dietary information in the form of plant microfossils 
[@henryCalculusSyria2008; @hardyStarchGranulesDental2009], 
proteins [@hendyProteomicCalculus2018; @warinnerDirectEvidenceMilk2014],
and other organic residues [@buckleyDentalCalculusCooking2014]. 
It has also proven to be able to retain and preserve dietary information for millenea,
providing detailed dietary information on past populations 
[@henryCalculusSyria2008; @jovanovicNeolithicCalculus2021; @taoWheatCalculus2020] 
and species [@henryNeanderthalCalculus2014; @hardyNeanderthalMedics2012].

<!-- more detailed background, comprehensible to scientists in related disciplines -->

Until recently, few studies directly investigated the presence of plant microremains
in the dental calculus of archaeological remains. The ability to extract phytoliths
from the dental calculus of archaeological fauna to investigate diet was first 
noted by Armitage [-@armitageExtractionIdentificationOpal1975], 
and later by Middleton and Rovner [-@middletonExtractionOpalPhytoliths1994],
and Fox and colleagues [-@foxPhytolithCalculus1996]. The ability to detect coca
consumption by looking at human dental calculus was explored by Klepinger and Kuhn
[-@klepingerCalculusCoca1977], and starches and phytoliths were extracted from 
human dental calculus by Cummings and Magennis [-@cummingsMayanCalculus1997].

In more recent years, the study of dental calculus has increased exponentially, 
and the wealth of information contained within the mineralised matrix has been, 
to a larger extent, acknowledged. The applications of dental calculus span a wide
variety of archaeological research areas, such as oral microbiome 
characterisation (including pathogens) through the analysis of DNA and proteins 
[@adlerSequencingAncientCalcified2013; @warinnerPathogensHostImmunity2014],
microbotanical remains [@henryCalculusSyria2008; @hardyStarchGranulesDental2009;
@mickleburghNewInsightsConsumption2012], 
other organic residues and proteins from dietary compounds 
[@buckleyDentalCalculusCooking2014; @hendyProteomicCalculus2018], 
and nicotine-use [@eerkensDentalCalculusSource2018]. Especially the extraction 
of starch granules has become a rich source of dietary
information, as starch granules have proven to preserve well within dental calculus
over a variety of geographical and temporal ranges 
[@pipernoStarchGrains2008; @henryNeanderthalCalculus2014; @taoWheatCalculus2020;
@jovanovicNeolithicCalculus2021].

<!-- starches from dental calculus -->
Despite this, our knowledge of dental calculus and the incorporation pathways of 
the various markers is limited [@radiniFoodMultiplePathways2017], as is our 
knowledge of information-loss caused by these pathways and the methods we use to 
extract and analyse dental calculus and make inferences on past diets. 
Extraction methods were tested 
by Tromp and colleagues [-@trompEDTACalculus2017], specifically regarding
decalcification using HCl or EDTA. 
The authors found significantly more starches with the EDTA extraction method 
than the HCl extraction method; however, as noted by the authors, comparisons 
involving archaeological calculus are problematic due to variability between and 
within individuals. Other studies have been conducted on modern populations 
[@leonardDentalCalculus2015] and primates [@powerChimpCalculus2015], both with
well-documented dietary intake, to reconcile our interpretations on the recovery 
of microremains (both phytoliths and starches) from dental calculus and how well
they represent actual dietary intake. These studies are justifiably limited, 
despite meticulous documentation and observation, due to unknown variables and 
uncertainty involved when studying living organisms. Dental calculus is a complex 
oral biofilm with a
multifactorial aetiology and variable formation rates both within and between
individuals [@jepsenCalculusRemoval2011].
Additionally, the effects of the many different methods
of starch processing, as well as post-depositional processes, are still being 
explored [@graneroStarchTaphonomy2020].

<!-- clearly stating the general problem being addressed by this particular study -->
<!-- Dental calculus research has its limitations, all of which I will solve with this
single method. -->
In this exploratory study, we use an oral biofilm model to investigate the 
retention of starch granules within dental calculus, in a controlled laboratory 
setting. Our main questions concern the representation of granules extracted
from the calculus compared to the actual intake. How much of the original diet is 
incorporated into the calculus, and how much is recovered? 
Is there differential loss of information from specific dietary markers that affects 
the obtained dietary information, and how does this affect the representation of diet from
extracted microremains? We find that a limited proportion of the starch input is actually 
retained in the calculus, as well as a shift in the size ratios of individual starch granules
that are incoporated into the calculus. We also show that the number of starch
granules that are incorporated increases as the size of the calculus deposit 
increases.

# Materials and Methods

```{r protocol-fig, fig.cap="Overview of experiment protocol"}
knitr::include_graphics(here("analysis/figures", "ExpL_protocol.png"))
```

<!-- remove sampling from protocol picture -->

## Biofilm formation

In this study we employ a multispecies oral biofilm model following a modified
protocol from Sissons and colleagues [-@sissonsMultistationDentalPlaque1991] and
Shellis [-@shellisSyntheticSalivaCultural1978]. The setup comprises a polypropylene 
24 deepwell PCR plate (KingFisher 97003510) with a lid containing 24 pegs (substrata), which 
is autoclaved at 120&deg;C, 1 bar overpressure, for 20 mins.

The artificial saliva (AS) is a modified version of the basal medium mucin (BMM)
described by Sissons and colleagues [-@sissonsMultistationDentalPlaque1991]. 
It contains 2.5 g/l partially purified mucin from porcine stomach (Type III, Sigma M1778), 
5 g/l trypticase peptone (Roth 2363.1), 10 g/l proteose peptone (Oxoid LP0085),
5 g/l yeast extract (BD 211921), 2.5 g/l KCl, 0.35 g/l NaCl, 1.8 mmol/l CaCl~2~, 
5.2 mmol/l Na~2~HPO~4~ [@sissonsMultistationDentalPlaque1991], 6.4 mmol/l NaHCO~3~ 
[@shellisSyntheticSalivaCultural1978], 2.5 mg/l haemin. This is subsequently 
adjusted to pH 7 with NaOH pellets and stirring, autoclaved (15 min, 120&deg;C, 
1 bar overpressure), and supplemented with 5.8 $\mu$mol/l menadione, 5 mmol/l urea, 
and 1 mmol/l arginine [@sissonsMultistationDentalPlaque1991].

<!-- add to protocols.io -->
Fresh whole saliva (WS) for inoculation was provided by a 31-year-old male donor 
with no history of caries, who abstained from oral hygiene for 24 hours, and no 
food was consumed two hours prior to donation. No antibiotics were taken up to six 
months prior to donation. 
The saliva was filtered through a sterilised (with bleach) nylon cloth to remove
particulates.
Substrata were inoculated with 1 ml/well of a two-fold dilution of WS in sterilised 
20% glycerine for four hours at 36&deg;C, to allow attachment of the
salivary pellicle and plaque-forming bacteria. After initial inoculation, the 
substrata were transferred to a new plate containing 1 ml/well AS and incubated 
at 36&deg;C, 30 rpm. The inoculation process was repeated on days 3 and 5. 
AS was partially refreshed once per day and fully refreshed every three days, 
throughout the experiment, by transferring the substrata to a new plate containing 
AS. To feed the bacteria, the substrata were transferred to a new plate, containing 
5% (w/v) sucrose, for six minutes twice daily, except on inoculation days 
(days 0, 3, and 5), where the samples only received one sucrose treatment after
inoculation.

Starch treatments were initiated on day 9 to avoid starch granule counts being 
affected by $\alpha$-amylase hydrolysis. An $\alpha$-amylase (EC 3.2.1.1) activity
assay was conducted to confirm that no amylase was present in the system (see 
Bartholdy et al. ???? <!-- ## [bartholdy] ## --> for more details). Starch treatments replaced sucrose treatments, occurring twice per day
for six minutes. The starch treatments involved transferring the substrata to a 
new plate containing a 0.25% (w/v) starch from potato (Roth 9441.1) solution, a 0.25% (w/v) starch from wheat (Sigma S5127) solution, and a 0.5% (w/v) mixture of equal
concentrations (w/v) wheat and potato. All starch solutions were created in a 5%
(w/v) sucrose solution. Before transferring biofilm samples to the starch treatments, 
the starch plates were agitated to keep the starches in suspension in the solutions, 
and during treatments, the rpm was increased to 60.

After 15 days, mineralisation was encouraged with a 
calcium phosphate monofluorophosphate urea (CPMU) solution containing 
20 mmol/l CaCl~2~, 12 mmol/l NaH~2~PO~4~, 5 mmol/l Na~2~PO~3~F, 500 mmol/l Urea
[@pearceConcomitantDepositionStrontium1987; @sissonsMultistationDentalPlaque1991], 
and (0.04 g/l MgCl). The substrata were submerged in 1 ml/well CPMU five times daily, 
every two hours, for six minutes. During the mineralisation period, starch treatments
were reduced to once per day after the five CPMU treatments. This cycle was repeated 
for 10 days until the end of the experiment on day 24 
(see \@ref(fig:protocol-fig) for an overview of protocol). More detailed protocols
are available on protocols.io<!--insert DOI-->.

All laboratory work was conducted in sterile conditions under a laminar flow hood 
to prevent starch and bacterial contamination. Control samples were included to 
detect starch contamination.

## Amylase activity detection

<!-- amylase protocol description -->

## Starch counts

Starch granule counts were divided into three size categories: small (<10 $\mu$m),
medium (10 -- 20 $\mu$m), and large (>20 $\mu$m). Counts for small wheat and
potato starches were combined for the mixed treatment samples, as it is not 
possible to distinguish the small starch granules between the two species.

### Treatment solutions

A 1 ml aliquot of each starch solution was taken, from which 10 $\mu$l was mounted 
on a microscope slide with an 18 x 18 mm coverslip, and counted under a light microscope
(Zeiss Axioscope A1). Counting starches on a full slide was not feasible, so three slide transects were 
counted (at 1/3, 1/2, and 2/3 of the slide), and the sample counts were extrapolated 
(see Supplementary material for more details).

### Extraction from samples

Extraction of starches from the calculus samples was performed by dissolving the
calculus in ethylenediaminetetraacetic acid (EDTA), as it has been shown to
have little to no effect on starch granules (especially native granules)
[@trompEDTACalculus2017; @lemoyneCalculusPretreatments2021; @modiCalculusMethodologies2020], 
and vortexing for 3 days until the sample was completely dissolved. 
Twenty $\mu$l of sample was mounted onto a slide with an 18x18 mm coverslip.
When dispersing the liquid (EDTA) sample on the slide, the sample was homogenised using the pipette. 
It was not possible to count all wheat starches
on a slide, so wheat and mix counts were extrapolated as described for the treatment
solutions.

## Statistical analysis

Statistical analysis was conducted in R versions 4.0.5 ("Shake and Throw") 
and 4.0.3 ("Bunny-Wunnies Freak Out") [@R-base].
<!-- R packages used -->
A one-way ANOVA with sample weight as the dependent variable (DV) and treatment 
as the grouping variable (GV) was conducted to explore the effect of the different 
starch treatments on biofilm growth.

To calculate size proportions, mean counts for each treatment were taken across 
both experimental plates for each treatment, resulting in a mean count for each
granule size category within each treatment.

Pearson's r was conducted on sample weight and total starch count, as well as sample
weight and starch count per mg calculus. The total count for each sample within a
treatment was standardised by z-score to account for the differences in magnitude
between the potato and wheat counts.

Scripts and data will be made available on osf.io upon publication.
Until then, they can be found on the first author's GitHub repository
(https://github.com/bbartholdy/byocStarch).

<!-- script and data availability: provide link (or doi?) to osf.io repository -->

# Results

All samples yielded sufficient biofilm growth and starch incorporation to be 
included in the analysis, resulting in a total of 48 biofilm samples (two plates of 24),
45 of which were used for analysis (three samples were set aside for separate analysis).
Most control samples contained no starch granules, while some contained negligible
quantities (supp. mat).

## No amylase activity detected in the model

<!-- insert amylase results -->

## Treatment type had no effect on biofilm growth

```{r anova-tbl}
knitr::kable(anova_tbl, 
             col.names = c("Treatment", "Mean", "SD", "Min", "Max"),
             caption = "Summary statistics for biofilm dry-weights (in mg) by treatment.")
```

A one-way ANOVA suggests that the type of starch used during the biofilm 
growth period had a minimal effect on the growth of the biofilm 
(expressed as total dry weight of the sample), F(`r paste(treat_df[1], treat_df[2], sep = ", ")`)
= `r treat_f.stat`, p = `r treat_p`. 
A summary of sample weights is available in Table \@ref(tab:anova-tbl).

## Starch counts

<!-- starch counts in the solutions -->

```{r solution-count-out}
count_sol_out <- count_sol %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(x / total, 0.1), ")")))

### adding the 'both' row to the data frame ###
count_sol_comb_out <- count_sol_comb %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(x / total, 0.1), ")")))

count_sol_out <- count_sol %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(x / total, 0.1), ")"))) %>%
  ungroup() %>%
  add_row(tibble_row(count_sol_comb_out), .before = 3)
```

```{r solution-count-tbl}
cnames <- c("Solution", "Starch", "Small (%)", "Medium (%)", "Large (%)", "Total (%)")
knitr::kable(count_sol_out, col.names = cnames, 
             caption = "Mean starch counts from solutions, including the
             proportional makeup of the different sizes of granules.")
```

<!-- starch counts in the samples -->

```{r sample-count}
count_sample <- count_sample %>%
  ungroup() %>%
  add_row(count_sample_comb, .after = 2)

# convert columns to contain mean count and percentage of total count
count_samp_per <- count_sample %>%
  mutate(across(c(s, m, l, total), 
                function(x) 
                  paste0(round(x), " (", scales::percent(signif(x / total, 3)), ")")))

# count_samp_comb_per <- count_sample_comb %>%
#   mutate(across(c(s, m, l, total), 
#                 function(x) 
#                   paste0(round(x), " (", scales::percent(x / total, 0.1), ")")))

# count_samp_comb_out <- tibble(count_samp_comb_per, count_sample_comb_sd, .name_repair = "unique") %>%
#   select(1,2,3,7,4,8,5,9,6,10) #%>%
  #add_row(tibble_row(count_sample_comb_out), .before = 3)

count_samp_out <- tibble(count_samp_per, count_sample_sd, .name_repair = "unique") %>%
  select(1,2,3,7,4,8,5,9,6,10) #%>%
  #add_row(tibble_row(count_sample_comb_out), .before = 3)
```

```{r sample-count-tbl}
cnames <- c("Treatment", "Starch", "Small (%)", "SD", "Medium (%)", "SD", 
            "Large (%)", "SD", "Total (%)", "SD")
knitr::kable(count_samp_out, col.names = cnames, 
             caption = "Mean starch counts extracted from samples with standard 
             deviation (SD), including the proportion of granule sizes of the total count.")
```

It was not possible to differentiate between potato and 
wheat starches smaller than ca. 10 $\mu$m,
small potato starches were counted as wheat starches in the mixed-treatment samples. This is reasonable given that the small potato granules make up an insignificant proportion of the 
total count of small granules within mixed-treatment solutions which are predominantly
wheat granules
(`r paste0(signif((count_sol$s[4] / sum(c(count_sol$s[3], count_sol$s[4]))) * 100, 3), "%")`).
<!-- count of small granules from wheat solution divided by the sum of small granules 
from the potato and wheat solutions -->

<!-- absolute solution counts compared to absolute sample counts. does exposure
to higher quantities of starches mean more incorporated starches? -->
The separate wheat and potato solutions were made with a 0.25% (w/v) starch
concentration, while the mixed-starch solution was made with 0.25% (w/v) of each
starch, with a total concentration of 0.50% (w/v).
The mixed-treatment had the highest absolute count of starch granules in solution,
while the biofilms exposed to the wheat solution preserved the greatest number of
granules. The potato treatment had the lowest absolute counts in both the solution 
and in the biofilm samples <!-- insert counts? -->
(Tables \@ref(tab:solution-count-tbl) and \@ref(tab:sample-count-tbl)).

<!-- look at proportions of incorporated starches in each sample before looking at overall means -->

### Proportion of available starches incorporated in samples

```{r sample-perc}
# convert proportions to percent labels
perc_comb <- prop_comb %>% 
  mutate(across(c(s, m, l, total), scales::percent, 0.001))

perc <- prop %>%
  mutate(across(c(s, m, l, total), scales::percent, 0.001)) %>% 
  add_row(tibble_row(perc_comb), .before = 3)
```

```{r sample-prop-tbl}
cnames <- c("Treatment", "Starch", "Small", "Medium", "Large", "Total")
knitr::kable(perc, col.names = cnames, 
             caption = "The mean percentage of starches from the solutions that 
              were incorported into the samples.")
```

The proportion of total starches from the solutions that were incorporated 
into the samples ranged from 
`r paste(min(perc$total), max(perc$total), sep = " to ")`,
with potato granules being more readily incorporated than wheat in both the
separated- and mixed-treatment samples
(Table \@ref(tab:sample-prop-tbl)). 
There is an inverse relationship between the absolute starch count in the solutions 
and the proportional incorporation of starches in the biofilm samples, i.e. potato 
had the lowest absolute count in solutions, but the highest proportional 
incorporation, and vice versa for the mixed treatment.

<!-- separated treatments -->

<!-- mixed treatments -->
Wheat incorporation was most affected in the mixed-treatment samples, with only
`r paste0(perc$total[2])` of the total available starches being incorporated into
the sample, compared to
`r paste0(perc$total[4])`
in the separated wheat treatment.

### Size ratios differ between solutions and samples

<!-- size proportions -->
<!-- pie charts -->

```{r sep-plot, fig.cap="Proportion (%) of sizes of (A) starch granules in the wheat solution (outer ring) and extracted from the wheat-treatment samples (inner ring), and (B) in the potato solution (outer ring) and extracted from the potato-treatment samples (inner ring). l = large, m = medium, s = small."}
#plot_grid(pl_wheat1, pl_potato1, labels = c("A", "B"), label_y = 0.1)
pl_wheat1 + pl_potato1 + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A")
```

<!-- change colour of medium starches in D to match C -->
```{r mix-plot, fig.cap="Proportion of sizes of (A) wheat granules in the mixed solution (outer ring) and extracted from the wheat-treatment samples (inner ring), and sizes of (B) potato granules in the solution (outer ring) and extracted from the potato-treatment samples (inner ring). l = large, m = medium, s = small."}
#plot_grid(pl_wheat2, pl_potato2, labels = c("C", "D"), label_y = 0.1)
pl_wheat2 + pl_potato2 + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A")
```
<!-- Combine to a single figure? -->

<!-- overall size trends -->

Overall, medium starch granules had a higher mean rate of incorporation 
(`r scales::percent(mean(prop$m, na.rm = T), 0.001)`) 
than small 
(`r scales::percent(mean(prop$s, na.rm = T), 0.001)`) 
and large 
(`r scales::percent(mean(prop$l, na.rm = T), 0.001)`) 
starch granules across all treatments, while large potato starches had the lowest 
rate of incorporation across all treatments.

<!-- separated and mixed treatments -->
The difference in incorporation between the size categories resulted in a change
in size ratios between the original starch solutions and the extracted samples.
Large potato granules (> 20 $\mu$m) were most affected, with a 
`r paste0(abs(diff_sep$potato[3]), "%")` 
decrease in relative abundance in the potato-only treatment, and a
`r paste0(abs(diff_mix$potato[3]), "%")` 
decrease in mixed treatments. Medium granules increased in relative abundance 
across all samples, while small granules decreased in wheat treatments and 
increased in potato treatments 
(Figures \@ref(fig:sep-plot) and \@ref(fig:mix-plot)).

<!-- Chi square test for size counts? -->

<!-- Proportion and size ratios are the most important points to be made -->

### Biofilm weight correlated positively with extracted starch counts

```{r cor, echo=FALSE}
# prepare cor output for reporting
cor_r <- signif(count_cor$estimate, 3)
cor_ci <- signif(count_cor$conf.int, 3)
cor_p <- if(count_cor$p.value > 0.001){
  signif(count_cor$p.value, 3)
} else {
  paste("< 0.001")
}

cor2_r <- signif(starch_cor$estimate, 3)
cor2_ci <- signif(starch_cor$conf.int, 3)
cor2_p <- if(starch_cor$p.value > 0.001){
  signif(starch_cor$p.value, 3)
} else {
  paste("< 0.001")
}
# Convert r correlation result to text
direct_cor <- ifelse(cor_r < 0, "negative", "positive")
strength_cor <- if(cor_r >= 0.8){
  "very strong"
} else if(cor_r < 0.8 & cor_r >= 0.6){
  "strong"
} else if(cor_r < 0.6 & cor_r >= 0.4){
  "moderate"
} else if(cor_r < 0.4 & cor_r >= 0.2){
  "weak"
} else{
  "very weak"
}
```

```{r cor-plot, warning=FALSE, fig.cap="Scatter plot of sample weight and standardised starch count by Z-score for seprated treatments."}
pl_cor
```
<!-- with Z-score standardisation -->

Pearson's *r* suggests a 
`r strength_cor` `r direct_cor` 
correlation between the total weight of the biofilms and the total starch count 
(standardised by z-score) extracted from the samples across treatments, 
*r* = `r cor_r`, 
90%CI[`r paste(cor_ci[1], cor_ci[2], sep = ", ")`],
p = `r cor_p`
(Figure \@ref(fig:cor-plot)).

The same test was applied to total biofilm weight and starch count per mg 
calculus (also standardised by z-score), resulting in a weak positive correlation, 
*r* = `r cor2_r`, 
90%CI[`r paste(cor2_ci[1], cor2_ci[2], sep = ", ")`],
p `r cor2_p`
(Figure \@ref(fig:cor-plot2)).

<!-- size ratios in the larger deposits vs the smaller deposits?
  Are larger starches more proportionately represented in larger deposits?
  -->

```{r cor-plot2, warning=FALSE, fig.cap="Scatter plot of sample weight in mg and standardised count of starch grains per mg calculus."}
pl_cor2
```

@@

# Acknowledgements {-}

<!-- If you're reading this, thank you for taking the time to read the source 
documents. You're awesome! -->

# References cited

### 2021-08-20 analysis/paper/results.Rmd
further development of manuscript
combine sections into single document

```{r setup-res, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
#opts_knit$set(root.dir = '/Users/username/path_to_project')
#corr_counts <- readr::read_csv(here("analysis/data/derived_data", "corr_counts.csv"))
#corr_comb <- readr::read_csv(here("analysis/data/derived_data", "corr_comb.csv"))
#sol_corr <- readr::read_csv(here("analysis/data/derived_data", "sol_corr.csv"))
#source(here("analysis/scripts/analysis.R"))
#source(here("analysis/scripts/tables.R"))
#source(here("analysis/scripts/figures.R"))
#source(c("tables.R", "figures.R", "analysis.R"))
#code to include R scripts in .Rmd: r knitr::spin_child("tables.R")
```

@@

<!-- exclude problematic samples: st1C5 wheat and st2C6 potato, and  st2C5 
Use more formal outlier detection -->

@@

```{r sep-plot, fig.cap="(A) Proportion (%) of sizes of starch granules in the wheat solution (outer ring) and extracted from the wheat-treatment samples (inner ring), and (B) in the potato solution (outer ring) and extracted from the potato-treatment samples (inner ring). l = large, m = medium, s = small."}
#plot_grid(pl_wheat1, pl_potato1, labels = c("A", "B"), label_y = 0.1)
pl_wheat1 + pl_potato1 + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A")
```

>>>>

@@

## No amylase activity detected in the model

<!-- insert amylase results -->

@@

```{r sep-plot, fig.cap="Proportion (%) of sizes of (A) starch granules in the wheat solution (outer ring) and extracted from the wheat-treatment samples (inner ring), and (B) in the potato solution (outer ring) and extracted from the potato-treatment samples (inner ring). l = large, m = medium, s = small."}
#plot_grid(pl_wheat1, pl_potato1, labels = c("A", "B"), label_y = 0.1)
pl_wheat1 + pl_potato1 + plot_layout(guides = "collect") + plot_annotation(tag_levels = "A")
```
